---
title: "Maudr Full Code and Notes"
author: "Lewis Ward"
format: html
---

# A package to simulate enzyme kinetics data.

# Packages (via pacman())

```{r}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(here, fs, readxl, writexl, tidyverse, magrittr, broom, gridExtra, ggpmisc, rlang, xfun)
```

# Ask on 27/11: does package need to load packages that the function uses? See stackoverflow thread. Perhaps user will be loading this in a new R studio environment -- is this relevant, are we assuming user has limited R knowledge? 
    # Can do this one of two ways: package::function() or @import (apparently this avoids conflicts?).

# Data Setup: 

```{r}
# Randomly assign substrate and inhibition type to each student
assignReactionConditions <- function(
		students, # list of students
		rxn_params, # table with reaction parameters
		time = c(.17, .33, .5, .66, .83, 1), # measurements timepoints (min)
		substr_conc = c(0,2,4,8,10,20,40,80,160), # substrate concentration (mM)
		cuv_vol = 0.003, # volume of cuvette (l)
		eps = 6220, # extinction coefficient
		enz_vol = 0.1, # volume of enzyme added (ml)
		save = TRUE){ # do you want to write the prepared set to disk?

	if(file.exists(here("data", "student_names.xlsx"))) {
		print("Loading file student_names.xlsx")
		students <- read_xlsx(here("data", "student_names.xlsx"))
	} else {
		print("The file student.xlsx was not found in the folder data or it is there under different name. Please double check its loaction and name and try again.")
	}

	if(file.exists(here("data", "reaction_parameters.xlsx"))) {
		print("Loading file reaction_parameters.xlsx")
		rxn_params <- read_xlsx(here("data", "reaction_parameters.xlsx"))
	} else {
		print("The file enzyme_parameters.xlsx was not found in the folder data or it is there under different name. Please double check its loaction and name and try again.")
	}

	if(!exists("students")) {
		load(here("data", "student_names.rda"))
		print("Example file with student names was loaded into the environment.")
	}

	if(!exists("rxn_params")) {
		load(here("data", "reaction_parameters.rda"))
		print("Example file with reaction parameters was loaded into the environment.")
	}

	seed  <- sample(100:999, 1) # Record the seed to recreate the data later if needed
	set.seed(seed)

	print(paste0("The seed value for this round is: ", seed, ". If save is set to FALSE, record this number for future reference."))

	students_rxn_params <- students %>%
		mutate(student_id = paste(toupper(student_no), toupper(first_name), sep = "_"),
					 rxn_substrate = sample(unique(rxn_params$rxn_substrate), nrow(students), replace = TRUE),
					 inhibition_actual = sample(unique(rxn_params$inhibition_actual)[-1], nrow(students), replace = TRUE)) %>%
		uncount(., 2, .id = "id") %>%
		mutate(inhibition_actual = ifelse(id == 2, "no_inhibition", inhibition_actual)) %>%
		select(-student_no, -first_name, -surname, -id) %>%
		left_join(., rxn_params, by = c("rxn_substrate", "inhibition_actual")) %>%
		mutate(substrate_conc = list(substr_conc)) # Add substrate concentrations as list-column

	if(save == TRUE){
		write_csv(x = students_rxn_params, file = here("output", paste0(Sys.Date(), "_", "students-rxn-params", "_seed_", seed, ".csv"))) # Save the file with the conditions for each student - useful for debugging and checking later on
		return(students_rxn_params)
	}
	else {
		return(students_rxn_params)
	}
}

assignReactionConditions()

```

# Data Assignment:

```{r}
# Function to calculate reaction rate V, with some jitter added to substrate concentration
calculateReactionRate <- function(Vmax, Km, substrate_conc) {
  V <- Vmax * ifelse(substrate_conc > 0, jitter(substrate_conc), substrate_conc) /
    (Km + ifelse(substrate_conc > 0, jitter(substrate_conc), substrate_conc))
  return(V)
}

# Function to calculate gradients for each reaction rate, with some jitter added to the enzyme volume
calculateGradient <- function(Vmax, Km, substrate_conc, enz_vol, cuv_vol, eps) {
  V <- calculateReactionRate(Vmax, Km, substrate_conc)
  gradient <- V * jitter(enz_vol) / 1e6 / cuv_vol * eps
  return(gradient)
}

# Function to fit a non-linear model to the Michaelis-Menten curve and estimate Km and Vmax
estimateKmVmax <- function(df) {
  nls(rxn_rate ~ estimated_Vmax * substrate_conc / (estimated_Km + substrate_conc),
      data = df,
      start = list(estimated_Km = 5, estimated_Vmax = 0.05)
  ) %>%
    tidy() %>%
    select(parameter = term, estimate) %>%
    mutate(estimate = round(estimate, 2)) %>%
    pivot_wider(names_from = parameter, values_from = estimate)
}
```

# Re-write from here

```{r}
# Function to process students' reaction parameters and estimate Km and Vmax
processReactionParameters <- function(students_rxn_params) {
  students_rxn_params %>%
    mutate(rxn_rate = pmap(select(., Vmax, Km, substrate_conc), calculateReactionRate)) %>%
    unnest(c(rxn_rate, substrate_conc)) %>%
    mutate(rxn_rate = round(rxn_rate, 2)) %>%
    group_by(student_id, rxn_substrate, inhibition_actual) %>%
    nest() %>%
    mutate(estimated_params = map(data, estimateKmVmax)) %>%
    select(-data) %>%
    mutate(estimated_params = bind_cols(inhibition_type = inhibition_actual, estimated_params)) %>%
    unnest(estimated_params) %>%
    ungroup() %>%
    select(-inhibition_actual) %>%
    group_by(student_id, rxn_substrate) %>%
    nest() %>%
    summarise(estimated_params = list(data)) %>%
    unnest(estimated_params)
}

# Function to generate absorbance vs time data files for each student
generateAbsVsTimeData <- function(student_id, df, cuv_vol, eps) {
  time <- c(0.17, 0.33, 0.5, 0.66, 0.83, 1)
  df %>%
    mutate(gradient = pmap(select(., Vmax, Km, substrate_conc), calculateGradient, enz_vol = 0.1, cuv_vol = cuv_vol, eps = eps),
           rxn_time = list(time)) %>%
    unnest(c(rxn_time)) %>%
    unnest(c(substrate_conc, gradient)) %>%
    mutate(absorbance = round(gradient * rxn_time, 3),
           rxn_condition = case_when(
             inhibition_actual == "no_inhibition" ~ "without_inhibitor",
             TRUE ~ "with_inhibitor"
           )) %>%
    relocate(rxn_condition, .after = inhibition_actual) %>%
    select(student_id, rxn_substrate, inhibition_actual, rxn_condition, substrate_conc, rxn_time, absorbance) %>%
    pivot_wider(names_from = substrate_conc_char, values_from = absorbance) %>%
    ## Add pivot_wider in for ease of student and lecturer parsing. 
    group_by(student_id) %>%
    nest() %>%
    mutate(
      filename = paste(here("output", paste0(student_id, "_data_new.xlsx"))),
      file_output = map2(filename, data, ~write_xlsx(.y, .x))
    ) %>%
    select(-data, -file_output)
}

generateAbsVsTimeData()

# Function to perform all data assignments (just to make workspace neater)
processedstudentdata <- function(students_rxn_params, cuv_vol = 0.003, eps = 6220) {
  # Calculate reaction parameters and estimate Km and Vmax
  estimated_params <- processReactionParameters(students_rxn_params)
  
  # Creation of absorbance vs time data for each student
  generateAbsVsTimeData(student_id = students_rxn_params$student_id, df = students_rxn_params, cuv_vol = cuv_vol, eps = eps)
  
  return(estimated_params)
}

processReactionParameters(students_rxn_params)


```


## Pivoting data - done though seems to have messed something up.
## Inclusion of student_ID in actual df rather than name - rectify the above before doing this. 

# Look at the useR base script
# Generate students_rxn_params first as an object (as a function) then run generateAbsvstimedata function on it (this function should be in useR script already). 

# Potential ideas:
## Maybe add EPS, etc in file?

# Prepare answer files

```{r}
### Functions to generate plots

Note that for the Lineweaver-Burke plot the x axis limits need to be dynamically calculated to include the x axis intercept for the non-inhibited data, but I don't know how to do it. Currently the limit is hard-coded to -0.05 but it won't be sufficient in all cases.

# Absorbance vs time plot
plotAbsVsTime <- function(df){
	df %>%
		mutate(gradient = pmap(select(., Vmax, Km, substrate_conc), calculateGradient), # Calculate of gradients for different substrate concentrations
					 rxn_time = list(time)) %>%
		unnest(c(rxn_time)) %>%
		unnest(c(substrate_conc, gradient)) %>% # Expand the dataframe for reaction rate and concentration
		mutate(absorbance = round(gradient * rxn_time, digits = 3), # Calculate absorbance
					 substrate_conc = fct_relevel(as.character(substrate_conc), levels = c("0", "10", "20", "40", "80", "160"))) %>%
		ggplot() + aes(x = rxn_time, y = absorbance, colour = substrate_conc) + geom_point() + geom_smooth(method = "lm", se = FALSE) + facet_wrap(~inhibition_actual) +
		labs(x = "Reaction time (min)",
				 y = "Absorbance (AU)",
				 title = "Absorbance vs reaction time plot",
				 colour = "Substrate concentration (mM)") +
		theme_minimal() +
		theme(legend.position = "bottom")
}

# Michaelis-Menten plot
plotMM <- function(df) {
	df %>%
		mutate(rxn_rate = pmap(select(., Vmax, Km, substrate_conc), calculateV)) %>%  # Calculate reaction rates
		unnest(c(substrate_conc, rxn_rate)) %>% # Expand the dataframe for reaction rate and concentration
		ggplot() + aes(x = substrate_conc, y = rxn_rate, colour = inhibition_actual) + geom_point() +
		labs(x = "Substrate concentration (mM)",
				 y = "Reaction rate (∆abs/min)",
				 title = "Michaelis-Menten plot",
				 colour = "Type of inhibition") +
		theme_minimal() +
		theme(legend.position = "bottom")
}

# Lineweaver-Burke plot
plotLB <- function(df) {
	df %>%
		mutate(rxn_rate = pmap(select(., Vmax, Km, substrate_conc), calculateV)) %>%  # Calculate of gradients for different reaction rates
		unnest(c(substrate_conc, rxn_rate)) %>% # Expand the dataframe for reaction rate and concentration
		filter(substrate_conc != 0) %>%
		mutate(substrate_conc_reciprocal = 1/substrate_conc,
					 rxn_rate_reciprocal = 1/rxn_rate) %>%
		ggplot() + aes(x = substrate_conc_reciprocal, y = rxn_rate_reciprocal, colour = inhibition_actual) + geom_point() + expand_limits(x = -0.05) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE) +
		labs(x = "Reciprocal substrate concentration (mM)",
				 y = "Reciprocal reaction rate",
				 title = "Lineweaver-Burke plot",
				 colour = "Type of inhibition") +
		theme_minimal() +
		theme(legend.position = "bottom") +
		stat_poly_eq(aes(label =  paste(after_stat(eq.label), after_stat(rr.label), sep = "*\", \"*")))
}

### Generate all the plots

...and add the table with estimated Km and Vmax for the final print

students_answers_data <- students_rxn_params %>%
	left_join(., estimated_Km_Vmax, by = c("student_id", "rxn_substrate")) %>%
	mutate(inhibition_actual = fct_relevel(inhibition_actual, "no_inhibition", after = 0L)) %>%
	group_by(student_id) %>%
	nest() %>%
	mutate(abs_vs_time_plot = map(data, plotAbsVsTime),
				 mm_plot = map(data, plotMM),
				 lb_plot = map(data, plotLB)) %>%
	unnest(data) %>%
	select(student_id, rxn_substrate, inhibition_actual, abs_vs_time_plot, mm_plot, lb_plot, estimated_params) %>%
	filter(inhibition_actual != "no_inhibition")

### Combine all the elements and print the answer file for each student

Answers for each students are saved on a single page PDF and then combined into a multi-page document for ease of printing.

createAnswerPDF <- function(student_id, rxn_substrate, inhibition_actual, abs_vs_time_plot, mm_plot, lb_plot, estimated_params){
	table1 <- tableGrob(tibble("info" = c("student", "substrate"), " " = c(student_id, rxn_substrate)), rows = NULL, cols = NULL)
	table2 <- tableGrob(unnest(estimated_params), rows = NULL)
	layout <- rbind(c(1,2), c(3,3), c(3,3), c(4,5), c(4,5))
	answers <- grid.arrange(table1, table2, abs_vs_time_plot, mm_plot, lb_plot, layout_matrix = layout)
	return(answers)
}

ggsave(plot = marrangeGrob(pmap(students_answers_data, createAnswerPDF), nrow=1, ncol=1), here("output", "SIB2004_answers_all_students.pdf"), width = 210, height = 297, units = "mm", dpi = "retina")

# Solution for separate files for each student
createAnswerPDF2 <- function(student_id, rxn_substrate, inhibition_actual, abs_vs_time_plot, mm_plot, lb_plot, estimated_params){
	table1 <- tableGrob(tibble("info" = c("student", "substrate"), " " = c(student_id, rxn_substrate)), rows = NULL, cols = NULL)
	table2 <- tableGrob(unnest(estimated_params), rows = NULL)
	layout <- rbind(c(1,2), c(3,3), c(3,3), c(4,5), c(4,5))
	answers <- grid.arrange(table1, table2, abs_vs_time_plot, mm_plot, lb_plot, layout_matrix = layout)
	ggsave(plot = answers, path = here("output"), filename = paste0(student_id, "_answers_new.pdf"), units = "mm", width = 210, height = 297, dpi = "retina")
}

pmap(students_answers_data, createAnswerPDF2)


```

# Look at user option as to which type of answer file they'd like (separate, together, both?)